Avoiding the Global Sort: \break A Faster Contour Tree Algorithm \footnote{The full updated version of this paper is available on the arXiv \cite{rs-mvrms-14}}

### 2016-03-22 paint-contour-tree.tex

\begin{figure}[h]\centering
    \includegraphics[width=.26\linewidth]{figs/spikey1}%
    \hspace{.35in}
    \includegraphics[width=.26\linewidth]{figs/spikey2}%
    \hspace{.4in}
    \includegraphics[width=.215\linewidth]{figs/spikeyTree}
    \caption{Two surfaces with different orderings of the maxima, but the same contour tree.}
    \label{fig:relative}
\end{figure}

@@

\begin{proof} Consider a call to $\rain(x,\NN)$. If the interface is empty, then all of $\NN$
is in $\wet(x,\NN)$, so $\NN$ is trivially extremum dominant. So suppose the interface
is non-empty and consists of $\phi_1, \phi_2, \ldots, \phi_k$ (as denoted in the procedure).
By repeated applications of \Clm{cut-int}, $\NN_{k+1}$ contains $\wet(x,\MM)$. 
Consider $\wet(x,\NN_{k+1})$. The interface must exactly be $\phi_1, \phi_2, \ldots, \phi_k$.
So the only dry vertices are those in the boundaries $B_1, B_2, \ldots, B_k$. But these
boundaries are maxima.
\end{proof}

@@

@@

\medskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\init(\MM)$}

@@

\medskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\build(\MM)$}

@@

\medskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\update(\stack)$}

@@

\medskip
\ignore{	
	Suppose $\mcol(h) = \{c\}$ (so $h$ is split). 
	\begin{compactenum}
		\item Connect $h$ (in $\reeb(\MM)$) to $\h(c)$ and the unique minimum corresponding to split $h$. 
		\item Delete $h$ from $T(c)$, set $\h(c) = h$.
		\item End procedure.
	\end{compactenum}
	\item Let $\mcol(h) = \{c_1, c_2\}$ (so $h$ is merge).
	\item Connect $h$ in $\reeb(\MM)$ to $\h(c_1)$ and $\h(c_2)$.
	\item Delete all copies of $h$ from $T(c_1)$ and $T(c_2)$.
	\item Perform union of colors $c_1$ and $c_2$ (denote merged color as $c$). Merge heaps
	to get $T(c) = T(c_1) \cup T(c_2)$.
	\item Set $\h(c) = h$.

@@

\InSoCGVer{
\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Some technical remarks}
\label{sec:techmarks}
The following is a continuation from the end of \Sec{basics}.

\technicalRemarks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Proof of Surgery}
\label{sec:proofOfSurgery}
\surgeryBody
That is, Theorem~\ref{thm:jordan}, which is proven below.
\proofofDisconnected

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Join and Split Trees of Extremum Dominant Complexes}
\label{sec:jstedm}
In this section we prove Theorem~\ref{thm:contour-tree} (of \Sec{extreme}), which states the equivalence 
of the contour tree on extremum dominant complexes and the critical join tree of Definition~\ref{def:criticalJoin}.
This proof requires some definitions and notation, specially
the notions of \emph{join} and \emph{split} trees, as given by~\cite{csa-cctad-03}.
Conventionally, all edges are directed from higher to lower function value.

\joinTreeSplitTreeMerge

We are now ready to prove Theorem~\ref{thm:contour-tree}.
\proofofContourJoinEquiv

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%reeb graph

\section{Proof of \Lem{runTimeUpper}}
\label{sec:runTimeProof}
\proofofRunTime

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\leafAssignPathDecomp

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\lowerBoundbyPathDecomp

>>>>

% \begin{figure}[h]\centering
%     \includegraphics[width=.26\linewidth]{figs/spikey1}%
%     \hspace{.35in}
%     \includegraphics[width=.26\linewidth]{figs/spikey2}%
%     \hspace{.4in}
%     \includegraphics[width=.215\linewidth]{figs/spikeyTree}
%     \caption{Two surfaces with different orderings of the maxima, but the same contour tree.}
%     \label{fig:relative}
% \end{figure}

\begin{figure}[h]\centering
    \includegraphics[width=.23\linewidth]{figs/spikey1}%
    \hspace{.37in}
    \includegraphics[width=.23\linewidth]{figs/spikey2}%
    \hspace{.41in}
    \includegraphics[width=.21\linewidth]{figs/spikeyTree}
    \caption{Two surfaces with different orderings of the maxima, but the same contour tree.}
    \label{fig:relative}
\end{figure}

@@

\begin{proof} Consider a call to $\rain(x,\NN)$. If the interface is empty, then all of $\NN$
is in $\wet(x,\NN)$, so $\NN$ is trivially extremum dominant. So suppose the interface
is non-empty and consists of $\phi_1, \ldots, \phi_k$ (as denoted in the procedure).
By repeated applications of \Clm{cut-int}, $\NN_{k+1}$ contains $\wet(x,\MM)$. 
Consider $\wet(x,\NN_{k+1})$. The interface is exactly $\phi_1, \ldots, \phi_k$.
So the only dry vertices are those in the boundaries $B_1, \ldots, B_k$. But these
boundaries are maxima.
\end{proof}

@@

An \emph{initial} painting also requires that the restriction $\chi:X \to [|X|]$ is a bijection.
\end{definition}

@@

\smallskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\init(\MM)$}

@@

\smallskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\build(\MM)$}

@@

\smallskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\update(\stack)$}

@@

%\medskip
\ignore{	
	Suppose $\mcol(h) = \{c\}$ (so $h$ is split). 
	\begin{compactenum}
		\item Connect $h$ (in $\reeb(\MM)$) to $\h(c)$ and the unique minimum corresponding to split $h$. 
		\item Delete $h$ from $T(c)$, set $\h(c) = h$.
		\item End procedure.
	\end{compactenum}
	\item Let $\mcol(h) = \{c_1, c_2\}$ (so $h$ is merge).
	\item Connect $h$ in $\reeb(\MM)$ to $\h(c_1)$ and $\h(c_2)$.
	\item Delete all copies of $h$ from $T(c_1)$ and $T(c_2)$.
	\item Perform union of colors $c_1$ and $c_2$ (denote merged color as $c$). Merge heaps
	to get $T(c) = T(c_1) \cup T(c_2)$.
	\item Set $\h(c) = h$.

@@

% % 
% % 
% % \InSoCGVer{
% % \appendix
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % \section{Some technical remarks}
% % \label{sec:techmarks}
% % The following is a continuation from the end of \Sec{basics}.
% % 
% % \technicalRemarks
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % \section{Proof of Surgery}
% % \label{sec:proofOfSurgery}
% % \surgeryBody
% % That is, Theorem~\ref{thm:jordan}, which is proven below.
% % \proofofDisconnected
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % \section{Join and Split Trees of Extremum Dominant Complexes}
% % \label{sec:jstedm}
% % In this section we prove Theorem~\ref{thm:contour-tree} (of \Sec{extreme}), which states the equivalence 
% % of the contour tree on extremum dominant complexes and the critical join tree of Definition~\ref{def:criticalJoin}.
% % This proof requires some definitions and notation, specially
% % the notions of \emph{join} and \emph{split} trees, as given by~\cite{csa-cctad-03}.
% % Conventionally, all edges are directed from higher to lower function value. 
% % 
% % \joinTreeSplitTreeMerge
% % 
% % We are now ready to prove Theorem~\ref{thm:contour-tree}.
% % \proofofContourJoinEquiv
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%reeb graph
% % 
% % \section{Proof of \Lem{runTimeUpper}}
% % \label{sec:runTimeProof}
% % \proofofRunTime
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % \leafAssignPathDecomp
% % 
% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% % \lowerBoundbyPathDecomp
% % 
% % }
% %
